CRUSE to Safe Cycling in Ireland

### 2023-09-11 cruse-paper.qmd
Put in new format

---
format:
  gfm: default
  pdf: default
title: "CRUSE to Safe Cycling in Ireland: An Open Source Methodology to Support Active Travel"
# in future: pdf!
editor: 
  markdown: 
    wrap: sentence
---

## 10th Transport Research Arena, TRA 2024, Dublin, Ireland

*Dr. Robin Lovelace (a), Dr. Joey Talbot (a), Dan Brennan (b), Dr. Suzanne Meade (b), Elaine Brick (c), and Peter Wright (d)*

a University of Leeds, Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom, Leeds LS2 9J, UK.

b Transport Infrastructure Ireland, Parkgate Business Centre, Parkgate Street, Dublin 8, D08 DK10, Ireland.

c AECOM, Unit 6, Galway Technology Park, Parkmore, Galway, Ireland.

d AECOM, Winslade House, Winslade Park, Manor Drive, Clyst St Mary, EXETER, EX5 1FY, UK.

## Abstract

Under the EU Road infrastructure safety management (RISM) directive, the National Road Safety Strategy (RSS), and the Climate Action Plan Transport Infrastructure Ireland (TII) has a remit for road safety and decarbonizing a predominantly road-based network in Ireland.
To address data needs for both safety and project evaluation on the National Road Network (NRN), the Cycle Route Uptake and Scenario Estimation (CRUSE) Tool was developed.
While cycling in Ireland represents only 3% of total modal share, with higher intensities in urban areas, the levels of cycling collisions are disproportionately high at 20% of all serious injuries and 7% of all fatalities.
If Ireland is to meet its climate and safety targets, data to establish baseline cycling levels and future cycling levels is needed.

@@

As countries, like Ireland, invest in cycling, the number of killed and seriously injured cyclists must reduce too.
The CRUSE Tool provides estimates of cycling potential and routing for each county in Ireland, and works in both urban and rural settings, to enable monitoring of cycling safety.
With growth in the E-bike market, the tool will help inform inter-urban and rural networks to support the transfer of trips to sustainable modes for longer journeys.
The CRUSE Tool methodology and findings are directly relevant to addressing the challenges and opportunities faced by other NRAs.
The datasets resulting from the project are open access and can be used by both non-experts and professionals.

**Keywords:** Cycling, Cycling Exposure, Open-source, Road Safety, Active Travel.

>>>>

---
title: "CRUSE to Safe Cycling in Ireland"
subtitle: An Open Source Methodology to Support Active Travel
author:
  - name: Dr. Robin Lovelace
    affiliations:
        address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
  - name: Dr. Joey Talbot
    affiliations:
      - name: University of Leeds
        address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
  - name: Dan Brennan
    affiliations:
      - name: Transport Infrastructure Ireland
        address: Parkgate Business Centre, Parkgate Street, Dublin 8, D08 DK10, Ireland
  - name: Dr. Suzanne Meade
    affiliations:
      - name: Transport Infrastructure Ireland
        address: Parkgate Business Centre, Parkgate Street, Dublin 8, D08 DK10, Ireland
  - name: Elaine Brick
    affiliations:
      - name: AECOM
        address: Unit 6, Galway Technology Park, Parkmore, Galway, Ireland
  - name: Peter Wright
    affiliations:
      - name: AECOM
        address: Winslade House, Winslade Park, Manor Drive, Clyst St Mary, EXETER, EX5 1FY, UK
abstract: |
  Under the EU Road infrastructure safety management (RISM) directive, the National Road Safety Strategy (RSS), and the Climate Action Plan Transport Infrastructure Ireland (TII) has a remit for road safety and decarbonizing a predominantly road-based network in Ireland.

  To address data needs for both safety and project evaluation on the National Road Network (NRN), the Cycle Route Uptake and Scenario Estimation (CRUSE) Tool was developed.
  While cycling in Ireland represents only 3% of total modal share, with higher intensities in urban areas, the levels of cycling collisions are disproportionately high at 20% of all serious injuries and 7% of all fatalities.
  If Ireland is to meet its climate and safety targets, data to establish baseline cycling levels and future cycling levels is needed.

@@

  As countries, like Ireland, invest in cycling, the number of killed and seriously injured cyclists must reduce too.
  The CRUSE Tool provides estimates of cycling potential and routing for each county in Ireland, and works in both urban and rural settings, to enable monitoring of cycling safety.
  With growth in the E-bike market, the tool will help inform inter-urban and rural networks to support the transfer of trips to sustainable modes for longer journeys.
  The CRUSE Tool methodology and findings are directly relevant to addressing the challenges and opportunities faced by other NRAs.
  The datasets resulting from the project are open access and can be used by both non-experts and professionals.
keywords: 
  - Cycling
  - Open-source
  - Road Safety
  - Active Travel
date: last-modified
bibliography: bibliography.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: super
# title: "CRUSE to Safe Cycling in Ireland: An Open Source Methodology to Support Active Travel"
# format: 
#   format:
#   elsevier-pdf:
#     keep-tex: true
#     journal:
#       name: Journal Name
#       formatting: preprint
#       model: 3p
#       cite-style: super
  # html:
  #   output-file: index.html
  # pdf:
  #   output-file: cruse-paper.pdf 
  # gfm:
  #   output-file: README.md
# affiliations:
#   - id: cmu
#     name: Carnegie Mellon University
#     city: Pittsburgh
#     state: PA
#   - id: chicago
#     name: University of Chicago
#     city: Chicago
#     state: IL
# number-sections: true
# editor: 
#   # render-on-save: true
#   markdown: 
#     wrap: sentence
# # Bibliography:
# bibliography: references.bib
# # APA style:
csl: apa.csl
---

## 10th Transport Research Arena, TRA 2024, Dublin, Ireland {.unnumbered}

{{< pagebreak >}}

**Keywords:** .

# Introduction

[@lovelace2017]
